### 2020-09-22 LaTeX/sections/paper.tex
resolved th text blues + added the numbers for arbitrum

For better or worst, blockchain technologies like Ethereum have dramatically lowered the barrier to entry for developing and deploying financial technology. New tokens have been launched with a few clicks of a user interface, and large investment infrastructures have been developed and deployed with little regulatory oversight. Blockchain exchange services allow order-based trading of digital currencies, tokens, and other digital assets. Such exchanges are a key component to blockchain-based economic activity.

For years, exchanges have operated either entirely off-chain by trusted operators or partially on-chain by semi-trusted operators. A fully on-chain exchange is feasible but considered too slow, at least for a public blockchain like Ethereum. We believe it is a very good time to do a deep dive into understanding precisely \textit{how slow} for the following reasons: (1) public blockchains are becoming faster (both in theory and slowly in practice) and the efficiency of an on-chain exchange will increase in time, (2) demand for on-chain trading is exemplified by the recent popularity of dealer-based (or quote-based) trading like Uniswap and Curve Finance, and (3) stablecoins have become a popular and allow on-chain trading with pricing in USD (or other government currency).

@@

\paragraph{Central exchanges (CEX).} Traditional financial markets (\eg NYSE and NASDAQ) use order-matching systems to arrange trades. An exchange will list one or more assets (stocks, bonds, derivatives, or more exotic securities) to be traded with each given its own orderbook priced in a currency (\eg USD). Exchanges for blockchain-based assets (also called cryptoassets by enthusiasts) can operate the same way, using a centralized exchange (CEX) design where a firm operates the platform as a trusted third party in every aspect: custodianship over assets/currency being traded, exchanging assets fairly, and offering the best possible price execution. Security breaches and fraud (\eg MtGox~\cite{TheHisto45:online}, QuadrigaCX~\cite{SEBIOrde83:online}, and many others) in centralized exchanges have become a common source of lost funds for users, while accusations of unfair trade execution have been levelled but are difficult to prove.

\paragraph{On-chain orderbooks.} For trades between two blockchain-based assets (\eg a digital asset priced with a cryptocurrency, stablecoin, or second digital asset), order matching can be performed `on-chain' by deploying the order-matching system either on a dedicated blockchain or inside a decentralized application (DApp; \textit{a.k.a.} smart contract). In this model, traders entrust their assets to an autonomously operating DApp with known source code instead of a third party custodian that can abscond with or lose the funds. The trading rules will operate as coded, clearing and settling can be guaranteed, and order submission is handled by the blockchain---a reasonably fair and transparent system (but see front-running below). Finally, anyone can create an on-chain orderbook for any asset (on the same chain) at any time. While they sound ideal, performance is a substantial issue and the main subject of this paper.

@@

\paragraph{On-chain dealers.} Another on-chain advantage is that other smart contracts, not just human users, can initiate trades, enabling broader decentralized finance (DeFi) applications. This has fuelled a resurgence in on-chain exchange but through a quote-driven design rather than an order-driven one. Automated market makers  (\eg Uniswap) have all the trust advantages of an on-chain orderbook, plus they are very efficient relative to an on-chain orderbook. The trade-off is that they operate as a dealer---the DApp exchanges assets from its own inventory. This inventory is loaded into the DApp by an investor who will not make money on the trades themselves, but hope for long term profit through trading fees. By contrast, an orderbook requires no upfront inventory and does not have to charge trading fees (but can). Finally, there is a complicated difference in their price dynamics (\eg market impact of a trade, slippage between the best bid/offer and actual average execution price, \etc)---deserving of an entire research paper to precisely define. We leave it as an assertion that with equal liquidity, orderbooks have more favourable price dynamics for traders.

\paragraph{Hybrid designs.} Before on-chain dealers became prominent in the late 2010s, the most popular design was hybrid order-driven exchanges with some trusted off-chain components and some on-chain functionality. Some (\eg EtherDelta) were envisioned as operating fully on-chain, but performance limitations drove developers to move key components, such as the order matching system, off-chain to a centralized database. A landscape of DEX designs exist: many avoiding taking custodianship of assets off-chain, and virtually all (for order-driven markets) operate the orderbook itself off-chain. A non-custodial DEX solves the big issue of a CEX---the operator stealing the funds---however trade execution is still not provably fair. A mitigation is to issue a proof of correct execution to the DApp (\eg Loopring) but these proofs have blindspots (discussed in section~\ref{sec:rollups}). \cm offers key advantages in this model.

@@

\paragraph{Blockchain limitations and solution.} While an orderbook is a ledger and blockchains provide a distributed ledger, it is not straightforward to drop a continuous-time orderbook onto a blockchain. An older 2014 paper~\cite{clark2014decentralizing} on the `Princeton prediction market'~\cite{Bra13} motivates our work. The authors observe the following limitations of on-chain continous orderbooks: block intervals are slow and not continuous, there is no support for accurate time-stamping, transactions can be dropped or reordered by miners, fast traders can react to submitted orders/cancelations when broadcast to network but not in a block and have their orders appear first (as examined in later work on front-running: \cite{eskandari2019sok,daian2019flash}).

\paragraph{Call markets.} The researchers propose using a call market instead of a continuous-time market~\cite{clark2014decentralizing}. Orders are collected and placed into the orderbook over a window of time (\eg 1 or more blocks), then the market is closed and the orders are processed in batch: the best bids are matched to the best asks in order. If the prices overlap, the miner keeps the difference (which they could extract anyways through front-running). Call markets largely side-step front-running attacks from other traders because reordering trades has no impact (discussed more in section~\ref{sec:front}). The paper does not include an implementation and was envisioned as an alt-coin (Ethereum was still in development in 2014) with market closing being done by the miners themselves as part of the blockchain logic.

@@

In designing \cm within Ethereum's gas model, performance is the main bottleneck. For a call market, processing all the trades and closing the market is the most time consuming step. The most critical design decision is the data structure for holding orders. While data structures are well studied for many languages, Solidity/EVM has its own unique aspects (\eg gas refunds, a relatively cheap mapping data structure, only partial support for object oriented programming) that creates difficulties in assessing which will perform best without actually deploying and evaluating each variant.

When closing a call market, the orders are examined in order: highest to lowest price for bids, and lowest to highest price for asks. It most circumstances, the market closing algorithm does not have to consider any deeper bids/asks from the list when choosing whether the current best bid and ask can be fulfilled. The only exceptions are in the case of a tie on price or a canceled order, both of which we return to later. For this reason, the ideal data structure for storing bids/asks is a \textit{priority queue} (see Table~\ref{tab:PQ_API}) where each order's priority is its price. Specifically, we use two PQs---one for bids where the highest price is the highest priority, and one for asks where the lowest price is the highest priority.

@@

\item \textbf{Heap with Dynamic Array.} A heap is a type of binary tree data structure that comes in two forms of a (i) Max-Heap and (ii) Min-Heap. All the nodes of tree are in a specific order and the root of the tree always represents the highest priority item of the data structure (the largest and smallest values in the Max-Heap and Min-Heap respectively). We implemented a priority queue with a heap that stores its data in a dynamically-sized array.

@@

\item \textbf{Heap with Mapping.} In the above implementations, the heap stores the entire order (as a struct) in the heap. In this variant, we store the order struct in a Solidity mapping and store the mapping's key in the heap.

@@

\item \textbf{Linked List with Mapping.} Finally, we try a variant of a linked list using a Solidity mapping. The value of the mapping is a struct with order data and the key of the next (and previous) node in the list. The contract stores the key of the first node (head) in the list.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{fig/random_insertion.png}
\caption{\footnotesize{Gas costs for enqueuing random items into five different priority queue variants. For the x-axis, a value of 10 indicates it is the 10th number entered in the priority queue. The y-axis is the cost of enqueue in gas.}  \label{fig:random_insertion}}
\end{figure}

@@

\subsubsection{Dequeue performance.} For each PQ variant storing 50 random integers, the \texttt{Dequeue()} function is iterated until the data structure is empty. The total gas cost for fully dequeuing the priority queue variants is outlined in Table~\ref{tab:PQUnitTests}. These tests are performed using the following Ethereum gas metrics: block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei.\footnote{EthStats (July 2020): \url{https://ethstats.net/}} Dequeuing removes data from the contract's storage. Recall this results in a gas refund. Based on our manual estimates (EVM does not expose the refund counter), every variant receives the maximum gas refund possible (\ie half the total cost of the transaction). In other words, each of them actually consumes twice the \texttt{gasUsed} amount in gas before the refund, however none of them are better or worst based on how much of a refund they generate.

\subsubsection{Discussion.} Based on enqueuing, the heap variants are the cheapest in terms of gas, while based on dequeuing, the link list variants are cheapest. This is in accordance with the theoretical worst-case time complexity for each. However, (i) the linked list variants are materially cheaper than the heap variants at dequeuing, and (ii) dequeuing in a call market must be done as a batch, whereas enqueuing is paid for one at a time by the trader submitting the order, and (iii) Ethereum will not permit more than hundreds of orders so the asymptotic behaviour is not a significant factor. For these reasons, we suggest using a linked list variant for this specific application. As it can be seen in Figure~\ref{fig:random_insertion}, the associated cost for inserting elements into a linked list PQ is significantly greater than the linked list with mapping, as each insertion cause the creation of a new contracts. Accordingly, we choose to implement the call market with the linked list with mapping. Overall this PQ balances a moderate gas cost for insertion (\ie order submission) with one for removal (\ie matching the orders).

@@

\subsection{Cost/Benefit of cleaning up after yourself}
\label{sec:LLvsLLMapping}
\label{sec:gasrefund}

@@

One consequence of a linked list is that a new contract is created for every node in the list. Beyond being expensive for adding new nodes (a cost that will be bared by the trader in a call market), it also leaves a large footprint on the Ethereum blockchain, especially if we leave the nodes on the blockchain in perpetuity (\ie we just update the head node of the list and leave the previous head `dangling.'). However in a PQ, nodes are only removed from the head of the list; thus the node contracts could be `destroyed' one by one using an extra operation, \texttt{SELFDESTRUCT}, in the \texttt{dequeue()} function. As shown in Table~\ref{tab:cleaning}, the refund from doing this outweighs to the cost of the extra computation: gas costs are reduced from 721K to 557K.  This suggests a general principle: cleaning up after yourself will pay for itself in gas refunds. Unfortunately, this is not universally true as shown by applying the same principle to the Linked List with Mapping.

Dequeuing in a linked list with mapping can be implemented in two ways. The simplest approach is process a node, update the head pointer, and leave the `removed' node's data behind in the mapping untouched (where it will never be referenced again). Alternatively we can call \texttt{DELETE} on each mapping entry once we are done with a node in the PQ. Deleting a storage variable is identical to setting a non-zero variable to zero (\texttt{SSTORE} 0) that costs 20,000 gas but with 15,000 refunded---a net positive gas cost~\cite{wood2014ethereum}. As it can be seen in the last two rows of Table~\ref{tab:cleaning}, leaving the data on chain is cheaper then cleaning it up.

The lesson here is that gas refunds incentivize developers to clean up storage variables they will not use again, but it is highly contextual as to whether it will pay for itself. Further the cap on the maximum refund means that refunds are not fully received for large cleanup operations (however removing the cap impacts the miners' incentives to include the transaction). This is a complicated and under-explored area of Ethereum in the research literature. For our own work, we strive to be good citizens of Ethereum and cleanup to the extent that we can---thus all PQs in Table~\ref{tab:PQUnitTests} implement some cleanup and we select Linked List with Mapping and \texttt{DELETE} for \cm.

@@

Another simplifying assumption was to implement a \textit{collateralized} call market. We assume all trades are between ETH and an ERC20 token, all orders are pre-funded in the contract with ETH (for bids) and tokens (for asks), and once ETH or tokens are committed to an order, they cannot be withdrawn until the market closes. We revisit alternatives to this design below. This ensures all executed orders clear and settle (\ie no defaults on payment or delivery).

@@

\cm is written in 225 lines (SLOC) of Solidity (including 3 external libraries),\footnote{Github: linked removed for anonymity} a high level programming language that is syntactically similar to Java~\cite{Ethereum41:online}. We tested it with the Mocha testing framework (tests included in the codebase), as deployed using Truffle on Ganache-CLI to obtain our performance metrics. Table~\ref{tab:callmarket_functions} represents the call market's primary operations.

@@

The main research question is how many orders can be processed in a single Ethereum transaction when the closing the call market, \textblue{using Ethereum today}. As our previous experiments indicated, the choice of priority queue (PQ) implementation is the main influence on performance (see Table~\ref{tab:worst_case_matching}). We implemented a generic call market that interfaces with a generic PQ (at its own contract address) and ran experiments for each PQ implementation. We looked at the \textit{worst-case} for performance which is when every submitted bid and ask is marketable (\ie will require fulfillment). In the first two columns, we determined the highest number of orders that can be processed in a single call to \texttt{CloseMarket()} and not exceed the the current Ethereum block gas limit of 11,741,495. Since Ethereum will become more efficient over time, we also were interested in how much gas it would cost to execute 1000 pairs of orders which is given in the third column. The fourth column indicates the cost to the trader of submitting a bid or ask --- since this costs will vary depending on how many orders are already submitted (recall Figure~\ref{fig:random_insertion}), we average the cost of 200 order submissions.

@@

We implement a very simple market where orders all have the same volume: 1 token. Traders submit multiple orders to increase volume. To extend orders with a volume field, it must first be determined if the token are non-divisible or divisible (and to what precision). In this case, \texttt{MarketClose()} needs look at multiple bids to fulfill a single ask (and vice-versa) and declare its trading rules on partial fills. We also simplify the market rules around ties on price: we execut them FIFO (breaking front-running resistance for ties on prices). A common trading rule (that does resist front-running) is to fill ties in proportion to their volume (\ie \textit{pro rata} allocation)\footnote{If Alice and Bob bid the same price for 100 tokens and 20 tokens respectively, and there are only 60 tokens left in marketable asks, Alice receives 50 and Bob 10.} however this approach is also contingent on the divisibility of tokens. Consider the following corner-case: 3 equally priced bids of 1 non-divisible token and 1 ask at the same price. There is not good option: (1) the bids could all be dropped (fair but not market efficient), (2) bids could be prioritized based on time as in \cm (front-running is viable, but in this corner case only), or (3) the bid could be randomly chosen (\cf~\cite{mavroudis2019libra}; blockchain `randomness' is generally deterministic and manipulatable by miners~\cite{bonneau2015random,buenz2017proofs} and counter-measures could take a few blocks to select~\cite{boneh2018verifiable}).

@@

 We want the market to run on a periodic schedule => it is difficult for market participants to call the \texttt{Close()} function periodically.

@@

%= = = = = = Miners Close the Market = = = = = = =%
\paragraph{Miners.} The difference between the best bid and ask prices is called the \textit{bid-ask spread}. In our design, when the trade occurs between the the highest bid (the highest amount a buyer is willing to pay for an asset) and lowest ask (the lowest amount a seller is willing to accept for an asset) orders, the bid-ask spread is paid to the miner. There are possibilities that (i) no trade occurs or (ii) the bid-ask spread is zero (\ie the best bid and best ask prices are identical). So there is enough economic incentive for the the miner to execute the \texttt{CloseMarket()} function and get refunded as the refund amount could be potentially higher than the bid-ask spread. 
% Miners will compete for it

@@

\paragraph{Next Traders.} Another solution pattern is to process certain number of bid and ask orders upon every execution of the \texttt{CloseMarket()} function rather than treating them as one substantial transaction. A market participant $P_i$ would process $n$ orders from the previous market (\texttt{CloseMarket(n)}) when sending new orders to the current market. This process continues until all the orders in the previous markets are processed. % incentive? => contribution to the community!?
% Chop it smaller, pushes it around -> could have orders in next market pay to close previous market -> not enough orders come in -> fall back to traditional close -> progressive close

@@

\paragraph{Contract Pays.} An alternative solution is to design the market such that the last person to submit an order calls the \texttt{CloseMarket()} function, but in contrast to a normal transaction (where the person initiating the transaction must pay the fee), here the CallMarket contract pays the cost for closing the market and matching the orders respectively. To enforce this design we can use Solidity function modifiers; every time a new order is submitted, a function modifier checks whether (i) the auction period has to end and/or (ii) the maximum number of total orders has reached. If any of these two conditions are met, the \texttt{CloseMarket()} will be called. Again, market participants are charged with certain amounts of fees every time they submit an order, these fees are accumulated in the CallMarket. Once the \texttt{CloseMarket()} is successfully executed and orders are matched, the contract transfers its funds to that person. Note that here the person must still have enough gas to cover the execution of the transaction as the funds will be only transferred after the transaction is fully executed. However, market participants are incentivized to do so as they may receive more ethers than they have spent.

@@

\paragraph{A Relayer.} Meta transactions enable users to execute Ethereum functions without paying the gas. Rather than spending gas, users sign their intended action using their private keys and broadcast it to the network with no cost. A third party process (\textit{a relayer}) then crafts the actual transaction on user's behalf, sends the transaction to the Ethereum blockchain, and charges the base contract with the associated fees (see Figure~\ref{fig:meta_tx}). The required gas to pay for the \texttt{Match()} function could be collected as fees. So market participants are charged with certain amounts of fees every time they submit an order, these fees are accumulated in the CallMarket contract and will be used to pay for executing the \texttt{CloseMarket()} function.

@@

% = = = = = = = = = = = =Transferring the Collateral back to the traders = = = = = = = = = = = = =  %
\subsection{Collateralization.}

>>>>

For better or worse, blockchain technologies like Ethereum have dramatically lowered the barrier to entry for developing and deploying financial technology. New tokens have been launched with a few clicks of a user interface, and large investment infrastructures have been developed and deployed with little regulatory oversight. Blockchain exchange services allow order-based trading of digital currencies, tokens, and other digital assets. Such exchanges are a key component to blockchain-based economic activity.

For years, exchanges have operated either entirely off-chain by trusted operators or partially on-chain by semi-trusted operators. A fully on-chain exchange is feasible but considered too slow, at least for a public blockchain like Ethereum. We believe it is a very good time to do a deep dive into understanding precisely \textit{how slow} for the following reasons: (1) public blockchains are becoming faster (both in theory and slowly in practice) and the efficiency of an on-chain exchange will increase in time, (2) demand for on-chain trading is exemplified by the recent popularity of dealer-based (or quote-based) trading like Uniswap and Curve Finance, and (3) stablecoins have become popular and allow on-chain trading with pricing in USD (or other government currency).

@@

\paragraph{Central Exchanges (CEX).} Traditional financial markets (\eg NYSE and NASDAQ) use order-matching systems to arrange trades. An exchange will list one or more assets (stocks, bonds, derivatives, or more exotic securities) to be traded with each given its own orderbook priced in a currency (\eg USD). Exchanges for blockchain-based assets (also called crypto assets by enthusiasts) can operate the same way, using a centralized exchange (CEX) design where a firm operates the platform as a trusted third party in every aspect: custodianship over assets/currency being traded, exchanging assets fairly, and offering the best possible price execution. Security breaches and fraud (\eg MtGox~\cite{TheHisto45:online}, QuadrigaCX~\cite{SEBIOrde83:online}, and many others) in centralized exchanges have become a common source of lost funds for users, while accusations of unfair trade execution have been levelled but are difficult to prove.

\paragraph{On-chain Orderbooks.} For trades between two blockchain-based assets (\eg a digital asset priced with a cryptocurrency, stablecoin, or second digital asset), order matching can be performed `on-chain' by deploying the order-matching system either on a dedicated blockchain or inside a decentralized application (DApp; \textit{a.k.a.} smart contract). In this model, traders entrust their assets to an autonomously operating DApp with known source code instead of a third party custodian that can abscond with or lose the funds. The trading rules will operate as coded, clearing and settling can be guaranteed, and order submission is handled by the blockchain---a reasonably fair and transparent system (but see front-running below). Finally, anyone can create an on-chain orderbook for any asset (on the same chain) at any time. While they sound ideal, performance is a substantial issue and the main subject of this paper.

@@

\paragraph{On-chain Dealers.} Another on-chain advantage is that other smart contracts, not just human users, can initiate trades, enabling broader decentralized finance (DeFi) applications. This has fuelled a resurgence in on-chain exchange but through a quote-driven design rather than an order-driven one. Automated market makers  (\eg Uniswap) have all the trust advantages of an on-chain orderbook, plus they are very efficient relative to an on-chain orderbook. The trade-off is that they operate as a dealer---the DApp exchanges assets from its own inventory. This inventory is loaded into the DApp by an investor who will not make money on the trades themselves, but hope for long term profit through trading fees. By contrast, an orderbook requires no upfront inventory and does not have to charge trading fees (but can). Finally, there is a complicated difference in their price dynamics (\eg market impact of a trade, slippage between the best bid/ask and actual average execution price, \etc)---deserving of an entire research paper to precisely define. We leave it as an assertion that with equal liquidity, orderbooks have more favourable price dynamics for traders.

\paragraph{Hybrid Designs.} Before on-chain dealers became prominent in the late 2010s, the most popular design was hybrid order-driven exchanges with some trusted off-chain components and some on-chain functionality. Some (\eg EtherDelta) were envisioned as operating fully on-chain, but performance limitations drove developers to move key components, such as the order matching system, off-chain to a centralized database. A landscape of DEX designs exist: many avoiding taking custodianship of assets off-chain, and virtually all (for order-driven markets) operate the orderbook itself off-chain. A non-custodial DEX solves the big issue of a CEX---the operator stealing the funds---however trade execution is still not provably fair. A mitigation is to issue a proof of correct execution to the DApp (\eg Loopring) but these proofs have blindspots (discussed in section~\ref{sec:rollups}). \cm offers key advantages in this model.

@@

\paragraph{Blockchain Limitations and Solution.} While an orderbook is a ledger and blockchains provide a distributed ledger, it is not straightforward to drop a continuous-time orderbook onto a blockchain. An older 2014 paper~\cite{clark2014decentralizing} on the `Princeton prediction market'~\cite{Bra13} motivates our work. The authors observe the following limitations of on-chain continous orderbooks: block intervals are slow and not continuous, there is no support for accurate time-stamping, transactions can be dropped or reordered by miners, fast traders can react to submitted orders/cancelations when broadcast to network but not in a block and have their orders appear first (as examined in later work on front-running: \cite{eskandari2019sok,daian2019flash}).

\paragraph{Call Markets.} The researchers propose using a call market instead of a continuous-time market~\cite{clark2014decentralizing}. Orders are collected and placed into the orderbook over a window of time (\eg 1 or more blocks), then the market is closed and the orders are processed in batch: the best bids are matched to the best asks in order. If the prices overlap, the miner keeps the difference (which they could extract anyways through front-running). Call markets largely side-step front-running attacks from other traders because reordering trades has no impact (discussed more in section~\ref{sec:front}). The paper does not include an implementation and was envisioned as an alt-coin (Ethereum was still in development in 2014) with market closing being done by the miners themselves as part of the blockchain logic.

@@

In designing \cm within Ethereum's gas model, performance is the main bottleneck. For a call market, processing all the trades and closing the market is the most time consuming step. The most critical design decision is the data structure for holding orders. While data structures are well studied for many languages, Solidity/EVM has its own unique aspects (\eg gas refunds, a relatively cheap mapping data structure, only partial support for object oriented programming) that create difficulties in assessing which will perform best without actually deploying and evaluating each variant.

When closing a call market, the orders are examined in order: highest to lowest price for bids, and lowest to highest price for asks. In most circumstances, the market closing algorithm does not have to consider any deeper bids/asks from the list when choosing whether the current best bid and ask can be fulfilled. The only exceptions are in the case of a tie on price or a canceled order, both of which we return to later. For this reason, the ideal data structure for storing bids/asks is a \textit{priority queue} (see Table~\ref{tab:PQ_API}) where each order's priority is its price. Specifically, we use two PQs---one for bids where the highest price is the highest priority, and one for asks where the lowest price is the highest priority.

@@

\item \textbf{Heap with Dynamic Array.} A heap is a type of binary tree data structure that comes in two forms of a (i) Max-Heap and (ii) Min-Heap. All the nodes of the tree are in a specific order and the root always represents the highest priority item of the data structure (the largest and smallest values in the Max-Heap and Min-Heap respectively). We implemented a priority queue with a heap that stores its data in a dynamically-sized array.

@@

\item \textbf{Heap with Mapping.} In the above implementations, the entire order is stored (as a struct) in the heap. In this variant, we store the order struct in a Solidity mapping and store the mapping keys in the heap.

@@

\item \textbf{Linked List with Mapping.} Finally, we try a variant of a linked list using a Solidity mapping. The value of the mapping is a struct with order data and the key of the next (and previous) node in the list. The contract stores the key of the first node (head) and last node (tail) in the list.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{fig/random_insertion2.png}
\caption{\footnotesize{Gas costs for enqueuing random numbers into five priority queue variants. For the x-axis, a value of 10 indicates it is the 10th number entered in the priority queue. The y-axis is the cost of enqueue in gas.}  \label{fig:random_insertion}}
\end{figure}

@@

\subsubsection{Dequeue Performance.} For each PQ variant storing 50 random integers, the \texttt{Dequeue()} function is iterated until the data structure is empty. The total gas cost for fully dequeuing the priority queue variants is outlined in Table~\ref{tab:PQUnitTests}. These tests are performed using the following Ethereum gas metrics: block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei.\footnote{EthStats (July 2020): \url{https://ethstats.net/}} Dequeuing removes data from the contract's storage. Recall this results in a gas refund. Based on our manual estimates (EVM does not expose the refund counter), every variant receives the maximum gas refund possible (\ie half the total cost of the transaction). In other words, each of them actually consumes twice the \texttt{gasUsed} amount in gas before the refund, however none of them are better or worse based on how much of a refund they generate.

\subsubsection{Discussion.} Based on enqueuing, the heap variants are the cheapest in terms of gas, while based on dequeuing, the link list variants are cheapest. This is in accordance with the theoretical worst-case time complexity for each. However, (i) the linked list variants are materially cheaper than the heap variants at dequeuing, and (ii) dequeuing in a call market must be done as a batch, whereas enqueuing is paid for one at a time by the trader submitting the order, and (iii) Ethereum will not permit more than hundreds of orders so the asymptotic behaviour is not a significant factor. For these reasons, we suggest using a linked list variant for this specific application. As it can be seen in Figure~\ref{fig:random_insertion}, the associated cost for inserting elements into a linked list PQ is significantly greater than the linked list with mapping, as each insertion causes the creation of a new contract. Accordingly, we choose to implement the call market with the linked list with mapping. Overall this PQ balances a moderate gas cost for insertion (\ie order submission) with one for removal (\ie matching the orders).

@@

\subsection{Cost/Benefit of Cleaning up After Yourself}
\label{sec:gasrefund}

@@

One consequence of a linked list is that a new contract is created for every node in the list. Beyond being expensive for adding new nodes (a cost that will be bared by the trader in a call market), it also leaves a large footprint on the Ethereum blockchain, especially if we leave the nodes on the blockchain in perpetuity (\ie we just update the head node of the list and leave the previous head `dangling.'). However in a PQ, nodes are only removed from the head of the list; thus the node contracts could be `destroyed' one by one using an extra operation, \texttt{SELFDESTRUCT}, in the \texttt{dequeue()} function. As shown in Table~\ref{tab:cleaning}, the refund from doing this outweighs to the cost of the extra computation: gas costs are reduced from 721K to 557K.  This suggests a general principle: cleaning up after yourself will pay for itself in gas refunds. Unfortunately, this is not universally true as shown by applying the same principle to the linked list with mapping.

Dequeuing in a linked list with mapping can be implemented in two ways. The simplest approach is to process a node, update the head pointer, and leave the `removed' node's data behind in the mapping untouched (where it will never be referenced again). Alternatively we can call \texttt{DELETE} on each mapping entry once we are done with a node in the PQ. Deleting a storage variable is identical to setting a non-zero variable to zero (\texttt{SSTORE} 0) that costs 20,000 gas but with 15,000 refunded---a net positive gas cost~\cite{wood2014ethereum}. As it can be seen in the last two rows of Table~\ref{tab:cleaning}, leaving the data on chain is cheaper than cleaning it up.

The lesson here is that gas refunds incentivize developers to clean up storage variables they will not use again, but it is highly contextual as to whether it will pay for itself. Further the cap on the maximum refund means that refunds are not fully received for large cleanup operations (however removing the cap impacts the miners' incentives to include the transaction). This is a complicated and under-explored area of Ethereum in the research literature. For our own work, we strive to be good citizens of Ethereum and clean up to the extent that we can---thus all PQs in Table~\ref{tab:PQUnitTests} implement some clean up and we select linked list with mapping and \texttt{DELETE} for \cm.

@@

Another simplifying assumption was to implement a \textit{collateralized} call market. We assume all trades are between ETH and an ERC20 token, all orders are pre-funded in the contract with ETH (for bids) and tokens (for asks), and once ETH or tokens are committed to an order, they cannot be withdrawn until the market closes. \textblue{We revisit alternatives to this design below}. This ensures all executed orders clear and settle (\ie no defaults on payment or delivery).

@@

\cm is written in 324 lines (SLOC) of Solidity~\footnote{Github: linked removed for anonymity}, a high level programming language that is syntactically similar to Java~\cite{Ethereum41:online}. We tested it with the Mocha testing framework (tests included in the codebase), as deployed using Truffle on Ganache-CLI to obtain our performance metrics. Table~\ref{tab:cm_functions} represents \cm's primary operations.

@@

The main research question is how many orders can be processed in a single Ethereum transaction when the closing the call market, using Ethereum today. As our previous experiments indicated, the choice of priority queue (PQ) implementation is the main influence on performance (see Table~\ref{tab:worst_case_matching}). We implemented a generic call market that interfaces with a generic PQ (at its own contract address) and ran experiments for each PQ implementation. We looked at the \textit{worst-case} for performance which is when every submitted bid and ask is marketable (\ie will require fulfillment). In the first two columns, we determined the highest number of orders that can be processed in a single call to the \texttt{closeMarket()} and not exceed the the current Ethereum block gas limit of 11,741,495. Since Ethereum will become more efficient over time, we also were interested in how much gas it would cost to execute 1000 pairs of orders which is given in the third column. The fourth column indicates the cost to the trader of submitting a bid or ask --- since this costs will vary depending on how many orders are already submitted (recall Figure~\ref{fig:random_insertion}), we average the cost of 200 order submissions.

@@

We implement a very simple market where orders all have the same volume: 1 token. Traders submit multiple orders to increase volume. To extend orders with a volume field, it must first be determined if the token are non-divisible or divisible (and to what precision). In this case, \texttt{closeMarket()} needs to look at multiple bids to fulfill a single ask (and vice-versa) and declare its trading rules on partial fills. We also simplify the market rules around ties on price: we execute them FIFO (breaking front-running resistance for ties on prices). A common trading rule (that does resist front-running) is to fill ties in proportion to their volume (\ie \textit{pro rata} allocation)\footnote{If Alice and Bob bid the same price for 100 tokens and 20 tokens respectively, and there are only 60 tokens left in marketable asks, Alice receives 50 and Bob 10.} however this approach is also contingent on the divisibility of tokens. Consider the following corner-case: 3 equally priced bids of 1 non-divisible token and 1 ask at the same price. There is not good option: (1) the bids could all be dropped (fair but not market efficient), (2) bids could be prioritized based on time as in \cm (front-running is viable, but in this corner case only), or (3) the bid could be randomly chosen (\cf~\cite{mavroudis2019libra}; blockchain `randomness' is generally deterministic and manipulatable by miners~\cite{bonneau2015random,buenz2017proofs} and counter-measures could take a few blocks to select~\cite{boneh2018verifiable}).

@@

 We want the market to run on a periodic schedule => it is difficult for market participants to call the \texttt{closeMarket()} function periodically.

@@

%= = = = = = Miners Close the Market = = = = = = =%
\paragraph{Miners.} The difference between the best bid and ask prices is called the \textit{bid-ask spread}. In our design, when the trade occurs between the the highest bid (the highest amount a buyer is willing to pay for an asset) and lowest ask (the lowest amount a seller is willing to accept for an asset) orders, the bid-ask spread is paid to the miner. There are possibilities that (i) no trade occurs or (ii) the bid-ask spread is zero (\ie the best bid and best ask prices are identical). So there is enough economic incentive for the the miner to execute the \texttt{closeMarket()} function and get refunded as the refund amount could be potentially higher than the bid-ask spread. 
% Miners will compete for it

@@

\paragraph{Next Traders.} Another solution pattern is to process certain number of bid and ask orders upon every execution of the \texttt{closeMarket()} function rather than treating them as one substantial transaction. A market participant $P_i$ would process $n$ orders from the previous market (\texttt{closeMarket(n)}) when sending new orders to the current market. This process continues until all the orders in the previous markets are processed. % incentive? => contribution to the community!?
% Chop it smaller, pushes it around -> could have orders in next market pay to close previous market -> not enough orders come in -> fall back to traditional close -> progressive close

@@

\paragraph{Contract Pays.} An alternative solution is to design the market such that the last person to submit an order calls the \texttt{closeMarket()}function, but in contrast to a normal transaction (where the person initiating the transaction must pay the fee), here the CallMarket contract pays the cost for closing the market and matching the orders respectively. To enforce this design we can use Solidity function modifiers; every time a new order is submitted, a function modifier checks whether (i) the auction period has to end and/or (ii) the maximum number of total orders has reached. If any of these two conditions are met, \texttt{closeMarket()} will be called. Again, market participants are charged with certain amounts of fees every time they submit an order, these fees are accumulated in the CallMarket. Once the  \texttt{closeMarket()} is successfully executed and orders are matched, the contract transfers its funds to that person. Note that here the person must still have enough gas to cover the execution of the transaction as the funds will be only transferred after the transaction is fully executed. However, market participants are incentivized to do so as they may receive more ethers than they have spent.

@@

\paragraph{A Relayer.} Meta transactions enable users to execute Ethereum functions without paying the gas. Rather than spending gas, users sign their intended action using their private keys and broadcast it to the network with no cost. A third party process (\textit{a relayer}) then crafts the actual transaction on user's behalf, sends the transaction to the Ethereum blockchain, and charges the base contract with the associated fees. The required gas to pay for the \texttt{matchOrders()} function could be collected as fees. So market participants are charged with certain amounts of fees every time they submit an order, these fees are accumulated in the CallMarket contract and will be used to pay for executing the \texttt{closeMarket()}function.

@@

%Arbitrum numbers: 
% Heap w/ dynamic array: matching 38 pairs, cost: 38,863
% Heap w/ dynamic array: matching 1000 pairs, cost: 38,851
% Heap w/ dynamic array: 2000 average submission: cost (averaged): 39,169 (we have the plot too, if needed)

% = = = = = = = = = = = =Transferring the Collateral back to the traders = = = = = = = = = = = = =  %
\subsection{Collateralization}
